\section{Introduction}

\IEEEPARstart{T}{he} ``memory wall'' is becoming an increasing challenge for the
computers of tomorrow. This ``memory wall'' occurs when the on-chip speed of
computations exceed the on-chip memory capacities, creating the situation where
computing circuitry has to lie idle waiting for instructions/data to arrive from
off-chip memory locations and be put into the on-chip cache. The cache is the
quickest, most expensive, and smallest memory resource in a computer, usually
residing as closely on-chip as possible to the computational units, like the
\emph{central processing unit} (CPU). This results in a gap of time needed for
fetching the correct instructions/data from off-chip memory onto the chip
requiring the instructions/data for computations. \emph{Prefetching} is a method
for reducing this gap.

The idea behind prefetching is to have a predictive module moving
instructions/data from the slower off-chip memory (like main memory) to the much
faster cache memory on-chip \emph{before} it is needed. Hence, when the
processor requires new instructions/data, they are already in the cache and can
be made use of within a small amount of on-chip clock cycles, not having to wait
for the slower off-chip clock cycles which are dominated by the time it takes
for one or more chips to communicate (like on a bus).

The purpose of this report is to analyze and compare a DCPT prefetcher with
Tagged Sequential prefetcher using the M5 simulator, with both prefetchers
utilizing a maximum of 8KiB. We implement a DCPT prefetcher, as this design
proves to be very effective.
\todo[inline]{Confirm if we can actually phrase the above sentence as is. Can we prove that it is very effective?}
The prefetcher is implemented according to the
algorithm presented in ``Storage Efficient Hardware Prefetching using Delta
Correlation Prediction Tables'' by Grannaes, Jahre, and Natvig~\cite{dcpt}. The
storage limitation makes it important to examine how the 8 KiB can be used most
efficiently.

\todo[inline]{We need to explain the scheme chapter in the introduction as well!}

\todo[inline]{About the introduction:
- Introduces the larger research area that the paper is
a part of\\
- Introduces the problem at hand\\
- Explains the scheme\\}
